Let’s Disco – Publishing person-level data as Linked Data

### 2015-03-17 paper/ISWC 2015 Tutorial/ISWC-2015-tutorial.tex
last fix

\begin{abstract}
The DDI-RDF Discovery Vocabulary (Disco) is an RDF Schema vocabulary that enables the representation of person-level data sets and related metadata.
It is based on DDI (Data Documentation Initiative) which is a metadata standard related to the observation and measurement of human activity.
Using Disco, such data sets which are often used in different domains like the social, ecomical and behavoural sciences can be published as Linked Data. 
However, for a meaningful and complete publication of these data sets, additional established RDF vocabularies (e.g. RDF Data Cube, DCAT, SKOS, PROV-O, XKOS, and PHDD) are reused to a large extend. 
This combination allows for expressing information on metadata of data collections, single data sets, relationships between them, different versions of them and related provenance information. 
In this tutorial, we show how Disco is interwoven with these vocabularies and how these connections can be used in order to publish complex and domain-specific data sets.
In different real world use cases, we demonstrate the adoption and inclusion of Disco in existing information systems. 
%Existing DDI instances are stored in form of XML documents which can be transformed into DDI-RDF and therefore exposed as Linked Data.
%Additionally, we present typical metadata constraints of DDI metadata as well as how to formulate them by different constraint languages (e.g. OWL 2) and how to actually validate them.

@@

\section{Motivation}
The DDI-RDF Discovery Vocabulary (Disco) \cite{Bosch2012} is an RDF Schema vocabulary which can be utilized to represent person-level data sets and related metadata as Linked Data. 
It is based on the metadata standard DDI (Data Documentation Initiative) which enables the representation of the observation and measurement of human activity and which is often used in domains like the social, ecomical and behavoural sciences. 
The metadata of such data can get highly complex by e.g. necessary provenance information and relationships to data aggregates, data collections, and the physical data sources.
Hence beside Disco, additional widely accepted and adopted RDF vocabularies (e.g. RDF Data Cube\footnote{\url{http://www.w3.org/TR/vocab-data-cube/}}, DCAT\footnote{\url{http://www.w3.org/TR/vocab-dcat/}}, SKOS\footnote{\url{http://www.w3.org/TR/skos-reference/}}, and PROV-O\footnote{\url{http://www.w3.org/TR/prov-o/}}) are reused to a large extend \cite{Bosch2013} in order to enable a full data documentation, which is necessary to represent such data in full extent. 
In this tutorial, an introduction to Disco and its inter-connections to these vocabularies is given. 
It is presented how these connections can be used in order to publish complex domain-specific data sets.
In different real world use cases and by building a data set serving as running example, the presenters of the tutorial demonstrate the adoption and inclusion of Disco in existing information systems.
%During practical exercises, participants will have the possibility to elaborate an RDF representation of the running example or their own data sets and to formulate typical queries.
Participants are encouraged to present their own data sets and use cases where Disco may be applied.

>>>>

\begin{abstract}
The DDI-RDF Discovery Vocabulary (Disco) is an RDF Schema vocabulary that enables the representation of person-level data sets and related metadata.
It is based on DDI (Data Documentation Initiative) which is a metadata standard related to the observation and measurement of human activity.
Using Disco, such data sets which are often used in different domains like the social, ecomical and behavioral sciences can be published as Linked Data. 
However, for a meaningful and complete publication of these data sets, additional established RDF vocabularies (e.g. RDF Data Cube, DCAT, SKOS, PROV-O, XKOS, and PHDD) are reused to a large extend. 
This combination allows for expressing information on metadata of data collections, single data sets, relationships between them, different versions of them and related provenance information. 
In this tutorial, we show how Disco is interwoven with these vocabularies and how these connections can be used in order to publish complex and domain-specific data sets.
In different real world use cases, we demonstrate the adoption and inclusion of Disco in existing information systems. 
%Existing DDI instances are stored in form of XML documents which can be transformed into DDI-RDF and therefore exposed as Linked Data.
%Additionally, we present typical metadata constraints of DDI metadata as well as how to formulate them by different constraint languages (e.g. OWL 2) and how to actually validate them.

@@

\section{Motivation}
The DDI-RDF Discovery Vocabulary (Disco) \cite{Bosch2012} is an RDF Schema vocabulary which can be utilized to represent person-level data sets and related metadata as Linked Data. 
It is based on the metadata standard DDI (Data Documentation Initiative) which enables the representation of the observation and measurement of human activity and which is often used in domains like the social, ecomical and behavioral sciences. 
The metadata of such data can get highly complex by e.g. necessary provenance information and relationships to data aggregates, data collections, and the physical data sources.
Hence beside Disco, additional widely accepted and adopted RDF vocabularies (e.g. RDF Data Cube\footnote{\url{http://www.w3.org/TR/vocab-data-cube/}}, DCAT\footnote{\url{http://www.w3.org/TR/vocab-dcat/}}, SKOS\footnote{\url{http://www.w3.org/TR/skos-reference/}}, and PROV-O\footnote{\url{http://www.w3.org/TR/prov-o/}}) are reused to a large extend \cite{Bosch2013} in order to enable a full data documentation, which is necessary to represent such data in full extent. 
In this tutorial, an introduction to Disco and its inter-connections to these vocabularies is given. 
It is presented how these connections can be used in order to publish complex domain-specific data sets.
In different real world use cases and by building a data set serving as running example, the presenters of the tutorial demonstrate the adoption and inclusion of Disco in existing information systems.
%During practical exercises, participants will have the possibility to elaborate an RDF representation of the running example or their own data sets and to formulate typical queries.
Participants are encouraged to present their own data sets and use cases where Disco may be applied.
